SST_patterns

### 2019-08-31 SST_patterns.Rmd
Extracting data from netCDFS

library(tidyverse)
library(lubridate)
library(ggpubr)
library(zoo)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
library(viridis)
library(ggrepel)
library(maptools)
library(sp)
library(geosphere)
library(PBSmapping)
library(scales)
# library(grid)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
# library(ncdf4) # This was used to process NetCDF files in an earlier version
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
```

@@

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData")
load("Data/OISST.RData")
load("Data/CMC.RData")
load("Data/MUR.RData")
```

@@

# Loading MUR data

```{r}
# MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
# MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
# MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
# MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
# MUR_Hout_Bay <- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")
# 
# MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Hout_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Sea_Point) %>%
#   dplyr::rename(site = station)
# 
# MUR_SST$date <- (ymd(MUR_SST$date))
# MUR_SST <- MUR_SST %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   drop_na()
# 
# match_func <- function(df){
#   match <- SACTN_US  %>%  
#   left_join(df, by = c("site", "date")) %>% 
#   na.trim()
#   return(match)
# }
# 
# MUR <- match_func(df = MUR_SST) %>%
#   drop_na()
# 
# # save(MUR, file = "Data/MUR.RData")
# 
# match_plot <- function(df){
#   plot1 <- df %>% 
# ggplot(aes(x = date, y = temp)) +
#   geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
#   geom_line() +
#   facet_wrap(~ site, nrow = 2) +
#   theme_bw()
#   return(plot1)
# }
# 
# MUR_plot <- match_plot(df = MUR)
# MUR_plot
```

# CMC data

```{r}
# CMC_Sea_Point <- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
# CMC_Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
# CMC_Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
# CMC_Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
# CMC_Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")
# 
# CMC_SST <- rbind(CMC_Lamberts_Bay,CMC_Hout_Bay, CMC_Port_Nolloth, CMC_Saldanha_Bay, CMC_Sea_Point) %>%
#   dplyr::rename(site = station)
# 
# CMC_SST$date <- (ymd(CMC_SST$date))
# CMC_SST <- CMC_SST %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   drop_na()
# 
# match_func <- function(df){
#   match <- SACTN_US  %>%  
#   left_join(df, by = c("site", "date")) %>% 
#   na.trim()
#   return(match)
# }
# 
# CMC <- match_func(df = CMC_SST) %>%
#   drop_na()
# 
# # save(CMC, file = "Data/CMC.RData")
# 
# CMC_plot <- match_plot(df = CMC)
# CMC_plot
```

# Loading the OISST

```{r}
# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
#             col.names = c("lon", "lat", "temp", "date"))
# 
# # Visualise the data
# # To explore the data I visualise the min temperatures along the South african coastline.
# 
# OISST %>%
#   filter(date == min(date)) %>%
#   ggplot(aes(x = lon, y = lat)) +
#   geom_raster(aes(fill = temp))
# 
# load("Data/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data 
# 
# unique_pixel <- OISST %>%
#   select(lon, lat) %>%
#   unique()
# 
# # Select nearest 1 pixels (k = 1)
# # Here I use knnx to find the closes 1 pixels to the insitu sites
# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
#                       query = as.matrix(site_list_sub[,5:6]), k = 1)
# 
# pixel_match <- unique_pixel[match_index,] %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   mutate(site = site_list_sub$site)
# 
# # Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
# OISST_match <- OISST %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   filter(combi %in% pixel_match$combi)
# # save(OISST_match, file = "Data/OISST_match.RData")
# 
# OISST_sites <- OISST_match %>%
#   left_join(pixel_match, by = c("combi", "lon", "lat")) #%>%
#   #dplyr::rename(temp_OISST =temp)
# 
# OISST_sites <- OISST_sites %>% 
#   dplyr::mutate(date = as.Date(date)) %>% 
#   drop_na()
# # save(OISST_sites, file = "Data/OISST_sites.RData")
# 
# ## Matching the Insitu data with the OISST SST data
# OISST <- match_func(df = OISST_sites)
# # save(OISST, file = "Data/OISST.RData")
```

@@

>>>>

library(tidyverse)
library(lubridate)
library(ggpubr)
library(zoo)
library(FNN)
library(scales)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
# library(ncdf4) # This was used to process NetCDF files in an earlier version
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ncdf4) # library for processing netCDFs
library(ncdf4)
library(data.table)
library(plyr)
```

@@

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData")
```

@@

@@

# Extrcting the MUR data

```{r}
ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
csvDir <- "/home/amieroh/Documents"

bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
                    CC = c(25, 35, 340, 355), # Canary Current
                   CalC = c(35, 45, 225, 240), # California Current
                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))

#          1         2         3         4
# 12345678901234567890123456789012345678901
# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc

read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
  coords <- bbox[,region]
    nc <- nc_open(ncFile)
    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
    fNameStem <-
      substr(ncFile, pathLen + 10, pathLen + 38)
    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
      round(4)
    dimnames(sst) <- list(lon = nc$dim$lon$vals,
                          lat = nc$dim$lat$vals)
    nc_close(nc)
    sst <-
      as.data.frame(melt(sst, value.name = "temp"), row.names = NULL) %>%
      mutate(t = ymd(fDate)) %>%
      na.omit()
    fwrite(sst,
           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
           append = TRUE, col.names = FALSE)
    rm(sst)
}

# the list of files
ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strtDate <- str_sub(ncList[1], start = 1, end = 8)
endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)

# apply the function
system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))
```

# Extracting the CMC data

```{r}
ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_global"
csvDir <- "/home/amieroh/Documents"

bbox <- data.frame(BenC = c(-35, -25, 15, 20), # Benguela Current
                    CC = c(25, 35, 340, 355), # Canary Current
                   CalC = c(35, 45, 225, 240), # California Current
                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))

#          1         2         3         4
# 12345678901234567890123456789012345678901
# 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc

read_nc <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
  coords <- bbox[,region]
    nc <- nc_open(ncFile)
    pathLen <- nchar(paste0(ncDir, "/", region)) + 1
    fNameStem <-
      substr(ncFile, pathLen + 10, pathLen + 38)
    fDate <- substr(ncFile, pathLen + 1, pathLen + 8)
    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
      round(4)
    dimnames(sst) <- list(lon = nc$dim$lon$vals,
                          lat = nc$dim$lat$vals)
    nc_close(nc)
    sst <-
      as.data.frame(melt(sst, value.name = "temp"), row.names = NULL) %>%
      mutate(t = ymd(fDate)) %>%
      na.omit()
    fwrite(sst,
           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
           append = TRUE, col.names = FALSE)
    rm(sst)
}

# the list of files
ncList <- list.files(path = ncDir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strtDate <- str_sub(ncList[1], start = 1, end = 8)
endDate <- str_sub(ncList[length(ncList)], start = 1, end = 8)

# apply the function
system.time(llply(ncList, read_nc, region = "BC", csvDir = csvDir, .parallel = TRUE))

ncDir <- "/Volumes/Challenger_Deep/OceanData/CMC0.2deg-CMC-L4-GLOB-v2.0/netCDF"
CMC.csv.dir <- "/Volumes/Challenger_Deep/spatial/processed/CMC0.2deg-CMC-L4-GLOB-v2.0/WBC/daily"

system.time(read_nc(ncDir, "AC", CMC.csv.dir))
```

# Extracting the OISST data

```{r}
bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
                   CC = c(25, 35, 340, 355), # Canary Current
                   CalC = c(35, 45, 225, 240), # California Current
                   HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
                   row.names = c("latmin", "latmax", "lonmin", "lonmax"))

OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
OISST.csv.dir <- "/home/amieroh/Documents/OISST_subset"

#          1         2
# 1234567890123456789012345
# avhrr-only-v2.19810901.nc

# function to extract the dims and data from OISST netCDFs
read_nc <- function(ncFile, region = region, csvDir = csvDir) {
  coords <- bbox[, region]
  nc <- nc_open(ncFile)
  pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
  fNameStem <-
    substr(ncFile, pathLen + 1, pathLen + 13)
  fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
  LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
  LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
  sst <- ncvar_get(nc,
                   varid = "sst",
                   start = c(LonIdx[1], LatIdx[1], 1, 1),
                   count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
    round(4)
  dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat = nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  sst <-
    as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
    mutate(t = ymd(fDate)) %>%
    na.omit()
  fwrite(sst,
         file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(sst)
}

# the list of files
ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strtDate <- str_sub(ncList[1], start = 15, end = 22)
endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)

# apply the function
system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
# system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
# system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
# system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))

```
